Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2024-09-25 Paper.qmd
added toc. tweaked axis labels. added code to simplify stat reporting.

---
title: "Child Signals of Need in Utila, Honduras: An Exploratory Study"
execute: 
  echo: false
  warning: false
format: html
lightbox: true
editor: visual
tbl-cap-location: bottom
bibliography: signaling.bib
---

@@

vars_c <- c(
  CaregiverAge = "Age (years)",
  EducationLevelYears = "Education level (years)",
  IncomeCategoryN = "Monthly household income (Lempira)",
  number_adults = "Adults in household",
  AdultsChildcare = "Adults who provide childcare"
)

@@

Child sadness, crying, and temper tantrums were common. Caregivers reported that `r sum(na.omit(modeldf$SadFreqN >= 8))` (`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%) children were sad at least once a week, `r sum(na.omit(modeldf$CryFreqN >= 8))` (`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%) cried at least more than once a week, and `r sum(na.omit(modeldf$TantrumFreqN >= 8))` (`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%) threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). See @fig-barplot_SignalFreq for plot showing signaling frequencies and @fig-signal_alluvial_plot for plot showing the number of individuals who engaged in specific combinations of signal frequencies..

Conflict between children and caretakers was also common, with `r sum(na.omit(modeldf$ConflictFreqN >= 8))` children (`r signif(sum((na.omit(modeldf$ConflictFreqN >= 8)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%) having conflicts with parents more than once a week. As with the frequencies of various types of signaling, many children (`r sum(na.omit(modeldf$ConflictFreqN == 0))` \[`r signif(sum((na.omit(modeldf$ConflictFreqN == 0)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%\]) were described as never having conflicts with parents. See @fig-barplot_ConflictFreq for plot showing the frequencies of conflict between caretakers and children.

The frequency of alloparenting effort was bimodally distributed, with `r sum(na.omit(modeldf$AlloparentingFreqN == 0))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN == 0)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `r sum(na.omit(modeldf$AlloparentingFreqN >= 30))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN >= 30)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to help with childcare at least once a day. See @fig-barplot_AlloparentingFreq for plot showing the frequencies child alloparenting effort.

::: {#fig-barplot_SignalFreq}
```{r}
barplot_SignalFreq
```

>>>>

---
title: "Child Signals of Need in Utila, Honduras: An Exploratory Study"
execute: 
  echo: false
  warning: false
format:
  html:
    toc: true
lightbox: true
editor: visual
tbl-cap-location: bottom
bibliography: signaling.bib
---

@@

vars_c <- c(
  CaregiverAge = "Caregiver age (years)",
  EducationLevelYears = "Caregiver education level (years)",
  IncomeCategoryN = "Monthly household income (Lempira)",
  number_adults = "Adults in household",
  AdultsChildcare = "Household adults who provide childcare"
)

@@

```{r}
fmt_sum <- function(v){
  v <- na.omit(v)
  n <- length(v)
  thesum <- sum(v)
  thepct <- 100*signif(thesum/n, 2)
  str_glue("{thesum} children ({thepct}%)")
}
```

Child sadness, crying, and temper tantrums were common. Caregivers reported that `{r} fmt_sum(modeldf$SadFreqN >= 8)`  were sad at least once a week, `{r} fmt_sum(modeldf$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(modeldf$TantrumFreqN >= 8)` threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily; sadness: `{r} fmt_sum(modeldf$SadFreqN >= 30)`; crying: `{r} fmt_sum(modeldf$CryFreqN >= 30)`;  and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN >= 30)`. However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(modeldf$SadFreqN == 0)`; crying; `{r} fmt_sum(modeldf$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN == 0)`. See @fig-barplot_SignalFreq and @fig-signal_alluvial_plot.

Conflict between children and caretakers was also common, with `{r} fmt_sum(modeldf$ConflictFreqN >= 8)` having conflicts with parents more than once a week. As with the frequencies of various types of signaling, `{r} fmt_sum(modeldf$ConflictFreqN == 0)` were described as never having conflicts with parents. See @fig-barplot_ConflictFreq.

The frequency of alloparenting effort was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` being reported to help with childcare at least once a day. See @fig-barplot_AlloparentingFreq.

::: {#fig-barplot_SignalFreq}
```{r}
barplot_SignalFreq + guides(fill = guide_legend(title = 'Times/month', nrow = 1, position = 'top', reverse = T))
```
